# Logistic Regression

At first, we approach with a logistic regression. 
```{r}
library(broom)
library(tidyverse)
library(caret)
library(MASS)
```
```{r}
train$Inflight.wifi.service = as.factor(train$Inflight.wifi.service)
train$Departure.Arrival.time.convenient = as.factor(train$Departure.Arrival.time.convenient)
train$Ease.of.Online.booking = as.factor(train$Ease.of.Online.booking) 
train$Gate.location = as.factor(train$Gate.location)
train$Food.and.drink = as.factor(train$Food.and.drink)
train$Online.boarding = as.factor(train$Online.boarding)
train$Seat.comfort = as.factor(train$Seat.comfort)
train$Inflight.entertainment = as.factor(train$Inflight.entertainment)
train$On.board.service = as.factor(train$On.board.service)
train$Leg.room.service = as.factor(train$Leg.room.service)
train$Baggage.handling = as.factor(train$Baggage.handling)
train$Checkin.service = as.factor(train$Checkin.service)
train$Inflight.service = as.factor(train$Inflight.service)
train$Cleanliness = as.factor(train$Cleanliness)
train$satisfaction  = as.factor(train$satisfaction )
```

```{r}
test$Inflight.wifi.service = as.factor(test$Inflight.wifi.service)
test$Departure.Arrival.time.convenient = as.factor(test$Departure.Arrival.time.convenient)
test$Ease.of.Online.booking = as.factor(test$Ease.of.Online.booking) 
test$Gate.location = as.factor(test$Gate.location)
test$Food.and.drink = as.factor(test$Food.and.drink)
test$Online.boarding = as.factor(test$Online.boarding)
test$Seat.comfort = as.factor(test$Seat.comfort)
test$Inflight.entertainment = as.factor(test$Inflight.entertainment)
test$On.board.service = as.factor(test$On.board.service)
test$Leg.room.service = as.factor(test$Leg.room.service)
test$Baggage.handling = as.factor(test$Baggage.handling)
test$Checkin.service = as.factor(test$Checkin.service)
test$Inflight.service = as.factor(test$Inflight.service)
test$Cleanliness = as.factor(test$Cleanliness)
test$satisfaction  = as.factor(test$satisfaction )
```


```{r}
set.seed(5293)
mod_logit1 <- glm(satisfaction ~ Gender + Customer.Type + Age + 
                 Type.of.Travel + Class + Flight.Distance + Inflight.wifi.service + 
                 Departure.Arrival.time.convenient + Ease.of.Online.booking + 
                 Gate.location + Food.and.drink + Online.boarding + Seat.comfort +
                 Inflight.entertainment + On.board.service + Leg.room.service +
                 Baggage.handling + Checkin.service + Inflight.service +
                 Cleanliness + Departure.Delay.in.Minutes + Arrival.Delay.in.Minutes , data = train, family = "binomial")
```
We can check the performance of the initial logistic regression model
```{r}
summary(mod_logit1)

```
we will check the effects of departure delay since it exhibits strange behavior
```{r}
#Check the effects of departure delay since it exhibits strange behavior
library(effects)
plot(Effect(focal.predictors = c("Departure.Delay.in.Minutes"), mod = mod_logit1))
```

```{r}
#Refit with significant variables
mod_logit2 <- glm(satisfaction ~ Gender + Customer.Type + Age + 
                 Type.of.Travel + Class + Flight.Distance + Inflight.wifi.service + 
                 Departure.Arrival.time.convenient + Ease.of.Online.booking + 
                 Gate.location + Food.and.drink + Online.boarding + Seat.comfort +
                 Inflight.entertainment + On.board.service + Leg.room.service +
                 Baggage.handling + Checkin.service + Inflight.service +
                 Cleanliness + Departure.Delay.in.Minutes  , data = train, family = "binomial")
```
We realized that the arrival delay and departure delay variable are seriously correlated. The details of this will be talked about in the reflection section. After removing the arrival delay feature, we find the depature delay variable behaves as what supposed
```{r}
summary(mod_logit2)
#After removing the arrival delay feature, we find the depature delay variable behaves as what supposed
```



Results of logistic Regression
```{r}
pred <- predict(mod_logit2, type = "response", newdata = test)
confusion_matrix <- table(test$satisfaction, pred>0.7)
row.names(confusion_matrix)<- c("neutral or dissatisfied", "satisfied")
noquote("Confusion Matrix")
confusion_matrix
accuracy=(confusion_matrix[1]+confusion_matrix[4])/25976
accuracy
```

From the variable importance information we see that the most important variables are type of travel, customer type and checkin service. 
```{r}
varImp(mod_logit2, scale = FALSE)

```
Due to the number of variables we have, we will not try to visualize the results, otherwise it would be very difficult to read. 




